---
title: "part2_chisquare_sydney"
author: "Sydney Rilum"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n) # arranges in decreasing order
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% #keep only these sites
  filter(tail %in% c("B","W")) # excludes observations without recorded tail condition
```

## Find counts and proportions of lizards with broken or whole tails at each site
```{r}
utst %>% 
  count(site, tail) # long format
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail) # wide format, easier to compare across groups

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% # shows percentage/proportion of counts
  janitor::adorn_pct_formatting(digits = 2) %>% # rounded to 2 digits
  janitor::adorn_ns() # also displays actual counts
```

## Convert to a contingency table

Use `column_to_rowname()`.
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site") #variable is site, site names become column names instead of its own column
```

## Chi-square

Are tail breakage and site related?

Are the proportions so different that we think we have enough evidence to reject the H(null) of independence, to allow us to conclude that there is a significant effect of site on tail breakage.
```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

p = 0.001456

Using a significance level of 0.05, yes we have sufficient evidence to reject the null hypothesis of independence, and retain alternative hypothesis that site and tail breakage are not independent.

OR

There is a significant association between site and lizard tail breakage.

OR

A significant effect of site on tail breakage.

OR

Site and tail breakage are not independent. 

**There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).**


















